README

### 2024-03-06 paper_elsevier/paper_elsevier.Rmd
robspack to rtorf

---
title: robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
author:
  - name: Sergio Ibarra-Espinosa
    email: sergio.ibarra-espinosa@noaa.gov
    affiliation: a,b
    correspondingauthor: true
    footnote: 1
  - name: Lei Hu
    email: lhu@noaa.gov
    affiliation: b
address:
  - code: a
    organization: Cooperative Institute for Research in Environmental Sciences, CU Boulder
    addressline: 216 UCB, University of Colorado Boulder campus
    city: Boulder
    state: Colorado
    postcode: 80309
    country: United States
  - code: b
    organization: NOAA Global Monitoring Laboratory
    addressline: 325 Broadway
    state: Colorado
    postcode: 80309
    country: United States
footnote:
  - code: 1
    text: "This is the first author footnote."
abstract: |
  In this study, we present a new open-source R package `robspack`, 
  to read, process, select, and plot NOAA Observation Package (ObsPack)
  data products. We use a methane ObsPack data product as an example
  in this code base, but it can be easily modified to analyze ObsPack
  products for other greenhouse gasses. The R package starts with 
  creating a catalog of  all ObsPack files in each product. It then 
  reads all files and creates one database. While reading each ObsPack 
  file, it extracts site elevation and time zone information from the 
  file header and calculates sampling altitude in meters above ground 
  level and local time for individual samping events. Finally, it 
  processes and selects observations for inverse modeling purposes. 
  This package imports functions from data.table R package, which 
  contains C bindings with parallel implementation via Open-MP 
  [@dt]. data.table is faster than other 
  Python, Julia and R implementations for data-science, 
  providing a strong basis for robspack. robspack provides functions 
  to perform these tasks in a transparent and efficient way, 
  supporting open-source communities in environmental sciences.
keywords: 
  - ObsPack
  - NOAA
  - Greenhouse gases
journal: "Environmental Modelling & Software"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

@@

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. 
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `robspack` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `robspack`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process consists in creating a 
summary of the ObsPack files, reading them in an iteration process, 
filtering, and generating another output and plots.

@@

To install `robspack`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment.

```{r install, eval = F}
remotes::install_github("ibarraespinosa/robspack")
library(robspack)
```

@@

`robspack` is a collection of function organized together to read and 
process ObsPack files [@masarie2014obspack]. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. In other to facilitate this task,
we are letting to the public the directory shown below. This directory 
includes a README file with all the required information and scripts 
generate process each category. The structure of the directory is:

https://github.com/ibarraespinosa/robspack/tree/main/rscripts

@@

```{r libs, include = F, message=F, warning=F}
library(robspack)
library(data.table)
```

@@

The `robspack` functions are shown in table 1.

Table 1. Functions and classes in `robspack`.

@@

# Conclusion
In this manuscript we presented an robskpack, an R package to read 
and process CH4 ObsPack GLOBALVIEW+ published by the Global 
Monitoring Laboratory (GML) from the National Oceanographic and 
Atmospheric Administration (NOAA). robspack reads the text data which 
have different headers and organizes them in a common format. Then, 
this software applies calculations to filter observations by time and 
space. Finally, this software generates receptors in a suitable format 
that allows it to run HYSPLIT and generate footprints. This software 
does not provide methods to run HYSPLIT, but the user can follow the 
site https://www.ready.noaa.gov/HYSPLIT.php.

>>>>

---
title: rtorf An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
author:
  - name: Sergio Ibarra-Espinosa
    email: sergio.ibarra-espinosa@noaa.gov
    affiliation: a,b
    correspondingauthor: true
    footnote: 1
  - name: Lei Hu
    email: lhu@noaa.gov
    affiliation: b
address:
  - code: a
    organization: Cooperative Institute for Research in Environmental Sciences, CU Boulder
    addressline: 216 UCB, University of Colorado Boulder campus
    city: Boulder
    state: Colorado
    postcode: 80309
    country: United States
  - code: b
    organization: NOAA Global Monitoring Laboratory
    addressline: 325 Broadway
    state: Colorado
    postcode: 80309
    country: United States
footnote:
  - code: 1
    text: "This is the first author footnote."
abstract: |
  In this study, we present a new open-source R package `rtorf`, 
  to read, process, select, and plot NOAA Observation Package (ObsPack)
  data products. We use a methane ObsPack data product as an example
  in this code base, but it can be easily modified to analyze ObsPack
  products for other greenhouse gasses. The R package starts with 
  creating a catalog of  all ObsPack files in each product. It then 
  reads all files and creates one database. While reading each ObsPack 
  file, it extracts site elevation and time zone information from the 
  file header and calculates sampling altitude in meters above ground 
  level and local time for individual samping events. Finally, it 
  processes and selects observations for inverse modeling purposes. 
  This package imports functions from data.table R package, which 
  contains C bindings with parallel implementation via Open-MP 
  [@dt]. data.table is faster than other 
  Python, Julia and R implementations for data-science, 
  providing a strong basis for rtorf. rtorf provides functions 
  to perform these tasks in a transparent and efficient way, 
  supporting open-source communities in environmental sciences.
keywords: 
  - ObsPack
  - NOAA
  - Greenhouse gases
journal: "Environmental Modelling & Software"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

@@

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. 
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `rtorf` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `rtorf`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process consists in creating a 
summary of the ObsPack files, reading them in an iteration process, 
filtering, and generating another output and plots.

@@

To install `rtorf`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment.

```{r install, eval = F}
remotes::install_github("ibarraespinosa/rtorf")
library(rtorf)
```

@@

`rtorf` is a collection of function organized together to read and 
process ObsPack files [@masarie2014obspack]. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. In other to facilitate this task,
we are letting to the public the directory shown below. This directory 
includes a README file with all the required information and scripts 
generate process each category. The structure of the directory is:

https://github.com/ibarraespinosa/rtorf/tree/main/rscripts

@@

```{r libs, include = F, message=F, warning=F}
library(rtorf)
library(data.table)
```

@@

The `rtorf` functions are shown in table 1.

Table 1. Functions and classes in `rtorf`.

@@

# Conclusion
In this manuscript we presented an robskpack, an R package to read 
and process CH4 ObsPack GLOBALVIEW+ published by the Global 
Monitoring Laboratory (GML) from the National Oceanographic and 
Atmospheric Administration (NOAA). rtorf reads the text data which 
have different headers and organizes them in a common format. Then, 
this software applies calculations to filter observations by time and 
space. Finally, this software generates receptors in a suitable format 
that allows it to run HYSPLIT and generate footprints. This software 
does not provide methods to run HYSPLIT, but the user can follow the 
site https://www.ready.noaa.gov/HYSPLIT.php.

### 2024-03-06 paper_elsevier/paper_elsevier.tex
robspack to rtorf

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{robspack An R package for processing atmospheric greenhouse gas
observations in NOAA's ObsPack}
    \author[a,b]{Sergio Ibarra-Espinosa%
  \corref{cor1}%
  \fnref{1}}
   \ead{sergio.ibarra-espinosa@noaa.gov} 
    \author[b]{Lei Hu%
  %
  }
   \ead{lhu@noaa.gov} 
      \affiliation[a]{
    organization={Cooperative Institute for Research in Environmental
Sciences, CU Boulder},addressline={216 UCB, University of Colorado
Boulder
campus},city={Boulder},postcode={80309},state={Colorado},country={United
States},}
    \affiliation[b]{
    organization={NOAA Global Monitoring Laboratory},addressline={325
Broadway},postcode={80309},state={Colorado},country={United States},}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{This is the first author footnote.}

  \begin{abstract}
  In this study, we present a new open-source R package
  \texttt{robspack}, to read, process, select, and plot NOAA Observation
  Package (ObsPack) data products. We use a methane ObsPack data product
  as an example in this code base, but it can be easily modified to
  analyze ObsPack products for other greenhouse gasses. The R package
  starts with creating a catalog of all ObsPack files in each product.
  It then reads all files and creates one database. While reading each
  ObsPack file, it extracts site elevation and time zone information
  from the file header and calculates sampling altitude in meters above
  ground level and local time for individual samping events. Finally, it
  processes and selects observations for inverse modeling purposes. This
  package imports functions from data.table R package, which contains C
  bindings with parallel implementation via Open-MP \citep{dt}.
  data.table is faster than other Python, Julia and R implementations
  for data-science, providing a strong basis for robspack. robspack
  provides functions to perform these tasks in a transparent and
  efficient way, supporting open-source communities in environmental
  sciences.
  \end{abstract}
    \begin{keyword}
    ObsPack \sep NOAA \sep 
    Greenhouse gases
  \end{keyword}

@@

The NOAA ObsPack data is delivered to the public as NetCDF and text
files. The structure of the files including descriptor fields depend on
the type of file. For instance, the metadata from aircrfats is different
than surface stations, but all the files include concentrations and
other critical fields. An important information is the scale of
measurement, which isXX. Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome. The \texttt{robspack} package provides the GHG
science and research community a transparent and efficient tool to
process ObsPack products for GHG modeling and analyses.

In this manuscript we present \texttt{robspack}, an R package to read,
process and plot NOAA ObsPack data. For this release, we are focued on
the \(CH_4\) ObsPack GLOBALVIEW+ product. The general process consists
in create a summary of the ObsPack files, reading them in an iteration
process, filter and generating another output and plots.

@@

To install \texttt{robspack}, the user must have installed the R package
remotes and run the following script. This process will install all the
required dependencies, such as data.table, \texttt{cptcity}, an R
package with more than 7000 color palettes, and \texttt{lubridate}, a
package to manage time and dates \citep{lu, cpt}. Then, we call the
libraries to load the function into the environment.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"ibarraespinosa/robspack"}\NormalTok{)}
\FunctionTok{library}\NormalTok{(robspack)}
\end{Highlighting}
\end{Shaded}

@@

\texttt{robspack} is a collection of function organized together to read
and process ObsPack files \citep{masarie2014obspack}. The general
process consists in create a summary of the ObsPack files, reading them
in an iteration process, filter and generating another output. In other
to facilitate this task, we are letting to the public the directory
shown below. This directory includes a README file with all the required
information and scripts generate process each category. The structure of
the directory is:

https://github.com/ibarraespinosa/robspack/tree/main/rscripts

@@

The \texttt{robspack} functions are shown in table 1.

Table 1. Functions and classes in \texttt{robspack}.

@@

In this manuscript we presented an robskpack, an R package to read and
process CH4 ObsPack GLOBALVIEW+ published by the Global Monitoring
Laboratory (GML) from the National Oceanographic and Atmospheric
Administration (NOAA). robspack reads the text data which have different
headers and organizes them in a common format. Then, this software
applies calculations to filter observations by time and space. Finally,
this software generates receptors in a suitable format that allows it to
run HYSPLIT and generate footprints. This software does not provide
methods to run HYSPLIT, but the user can follow the site
https://www.ready.noaa.gov/HYSPLIT.php.

>>>>

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rtorf An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{rtorf An R package for processing atmospheric greenhouse gas
observations in NOAA's ObsPack}
    \author[a,b]{Sergio Ibarra-Espinosa%
  \corref{cor1}%
  \fnref{1}}
   \ead{sergio.ibarra-espinosa@noaa.gov} 
    \author[b]{Lei Hu%
  %
  }
   \ead{lhu@noaa.gov} 
      \affiliation[a]{
    organization={Cooperative Institute for Research in Environmental
Sciences, CU Boulder},addressline={216 UCB, University of Colorado
Boulder
campus},city={Boulder},postcode={80309},state={Colorado},country={United
States},}
    \affiliation[b]{
    organization={NOAA Global Monitoring Laboratory},addressline={325
Broadway},postcode={80309},state={Colorado},country={United States},}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{This is the first author footnote.}

  \begin{abstract}
  In this study, we present a new open-source R package
  \texttt{rtorf}, to read, process, select, and plot NOAA Observation
  Package (ObsPack) data products. We use a methane ObsPack data product
  as an example in this code base, but it can be easily modified to
  analyze ObsPack products for other greenhouse gasses. The R package
  starts with creating a catalog of all ObsPack files in each product.
  It then reads all files and creates one database. While reading each
  ObsPack file, it extracts site elevation and time zone information
  from the file header and calculates sampling altitude in meters above
  ground level and local time for individual samping events. Finally, it
  processes and selects observations for inverse modeling purposes. This
  package imports functions from data.table R package, which contains C
  bindings with parallel implementation via Open-MP \citep{dt}.
  data.table is faster than other Python, Julia and R implementations
  for data-science, providing a strong basis for rtorf. rtorf
  provides functions to perform these tasks in a transparent and
  efficient way, supporting open-source communities in environmental
  sciences.
  \end{abstract}
    \begin{keyword}
    ObsPack \sep NOAA \sep 
    Greenhouse gases
  \end{keyword}

@@

The NOAA ObsPack data is delivered to the public as NetCDF and text
files. The structure of the files including descriptor fields depend on
the type of file. For instance, the metadata from aircrfats is different
than surface stations, but all the files include concentrations and
other critical fields. An important information is the scale of
measurement, which isXX. Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome. The \texttt{rtorf} package provides the GHG
science and research community a transparent and efficient tool to
process ObsPack products for GHG modeling and analyses.

In this manuscript we present \texttt{rtorf}, an R package to read,
process and plot NOAA ObsPack data. For this release, we are focued on
the \(CH_4\) ObsPack GLOBALVIEW+ product. The general process consists
in create a summary of the ObsPack files, reading them in an iteration
process, filter and generating another output and plots.

@@

To install \texttt{rtorf}, the user must have installed the R package
remotes and run the following script. This process will install all the
required dependencies, such as data.table, \texttt{cptcity}, an R
package with more than 7000 color palettes, and \texttt{lubridate}, a
package to manage time and dates \citep{lu, cpt}. Then, we call the
libraries to load the function into the environment.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"ibarraespinosa/rtorf"}\NormalTok{)}
\FunctionTok{library}\NormalTok{(rtorf)}
\end{Highlighting}
\end{Shaded}

@@

\texttt{rtorf} is a collection of function organized together to read
and process ObsPack files \citep{masarie2014obspack}. The general
process consists in create a summary of the ObsPack files, reading them
in an iteration process, filter and generating another output. In other
to facilitate this task, we are letting to the public the directory
shown below. This directory includes a README file with all the required
information and scripts generate process each category. The structure of
the directory is:

https://github.com/ibarraespinosa/rtorf/tree/main/rscripts

@@

The \texttt{rtorf} functions are shown in table 1.

Table 1. Functions and classes in \texttt{rtorf}.

@@

In this manuscript we presented an robskpack, an R package to read and
process CH4 ObsPack GLOBALVIEW+ published by the Global Monitoring
Laboratory (GML) from the National Oceanographic and Atmospheric
Administration (NOAA). rtorf reads the text data which have different
headers and organizes them in a common format. Then, this software
applies calculations to filter observations by time and space. Finally,
this software generates receptors in a suitable format that allows it to
run HYSPLIT and generate footprints. This software does not provide
methods to run HYSPLIT, but the user can follow the site
https://www.ready.noaa.gov/HYSPLIT.php.

### 2024-03-06 rscripts/README.Rmd
robspack to rtorf

The directory `rscripts` contains the required files
to generate the list of receptors using `robspack`
The scripts will run from a defined working directory

>>>>

The directory `rscripts` contains the required files
to generate the list of receptors using `rtorf`
The scripts will run from a defined working directory
